Data Characteristics
The data used has been cleaned and rebased at an appropriate time period to allow for a better comparison between the nations, removing the influence of other factors such as currency exchange.
Energy price cpi - the average price of energy given in monthly increments, rebased at 2015=100 to aid in comparison between the three nations.

Headline cpi - the monthly all-items, non-adjusted consumer price index, rebased at 2002=100 to aid in comparison between the three nations.

Domestic consumption - the total domestic consumption of energy, given in quarterly increments, rebased at 2015=100 to aid in comparison between the three nations.

Energy supply - the type of energy production method used within the constituent nations.

Scatter Plots
CPI Energy Price vs Headline CPI
This scatter plot shows the relationship between the energy price cpi and the headline cpi for the three nations. Each point represents a monthly observation, with the x-axis representing the energy price cpi and the y-axis representing the headline cpi. 
The plot indicates a positive correlation between the two variables, suggesting that as energy prices increase, headline inflation also tends to increase.
General positive direction, although the year 2022 has some variation.

CPI Energy vs CPI Headline Slope
This plot is an amalgamation of the previous scatter plots for all the nations with the addition of the slope, which shows the corresponding change in headline CPI for every one unit rise in energy price.
The United States and Canada have similar slopes, which could roughly be expected due to their geographic location and economic structure.
The United Kingdom has a lower slope, indicating that headline CPI absorbed less of the energy shock per unit.
High R^2 values for each of the nations which need to be further interrogated. 

CPI Headline vs Con
This scatter plot illustrates the relationship between the headline cpi and domestic consumption. Each point represents a quarterly observation, with the x-axis representing the headline cpi and the y-axis representing domestic consumption.
The slope indicates the change in domestic consumption per one unit change in headline CPI.
Canada had the highest slope which was unexpected, meaning that there was strong growth in domestic consumption alongside the growth in headline CPI.
The United States had a nearly 1:1 relationship between the two variables, which was similarly unexpected.
The United Kingdom fared worse than both the US and Canada, with a much lower slope, although this was not unexpected.

CPI Headline vs Con EQUAL
Gives the same scatter plots as CPI Headline vs Con, but with only the data for years in which all three nations have data available for.
In the equalised version Canada and the United States have similar slopes near the 1:1 ratio, whilst the United Kingdom's remained relatively low at 0.664, showing that there was a meaningful difference in the growth between the two variables.
This time period is more recent and shows changes between the non-equalised version, it may be useful to include economic growth rates within the model to try and control for that factor.

Correlation Maps
CPI Energy vs CPI Headline Correlation
Shows the correlation between the energy price cpi and the headline cpi for each of the three nations. The correlation coefficient is calculated to quantify the strength and direction of the relationship between the two variables. A positive correlation coefficient indicates that as energy prices increase, headline inflation also tends to increase.
All three nations have a correlation close to one in the year 2022, showing that during that year there is an almost perfect correlation between the headline CPI and Energy CPI.
The United Kingdom and Canada have a higher median correlation value than the United States, indicating that they are generally more affected by energy price swings transmitting into headline CPI than the United States, which was to be expected.

CPI energy vs CPI Headline LAG
Shows how strongly a change in the energy CPI at a given time period predicts a change in the headline CPI.
The positive values shown indicate that energy CPI leads the headline CPI, demonstrating that changes in energy prices can be a predictor of future changes in headline inflation.
These variables have been tested to see if they are statistically significant, with the results showing that the relationship between energy prices and headline inflation is indeed significant for most of the measured values in Canada and the United Kingdom.
The United States is the only nation to not have many significant variables which warrants further investigation into their policy strategies at the time and how their economy reacted to the shock.

Energy CPI vs CPI Headline Individual
Shows the same as the CPI Energy vs CPI Headline, but with the three countries separated out instead of within the same graph for easier reference and identification of any significant changes and events present within the graph.

Energy Production All
Shows the methods of energy production over time between the three nations, showing how these methods changed and the percentage values of the different methods used.
It shows the total amount of energy produced within the economies and what share of it was produced by which type in a given year, with 2022 specifically shown as it's the year of the examined energy crisis.

Correlation Analysis
A first order effect does exist between the energy prices and the headline inflation rate as the correlation between the two variables is positive and relatively large.
The slope when comparing the two variables is similarly positive and relatively high, showing that there is a positive first order effect between the two variables, where one increases, so does the other.
The Simpson's Paradox is certainly worth considering, especially as the three countries do not share the same systems of governance or perhaps response to the crisis, an immediate example would be the United Kingdom's usage of an energy price cap. This should be included in the future as a comparison and to see whether the price cap was reached.
The slope of the United Kingdom is much lower than the two other nations, indicating that the price cap may have an effect or some other UK specific factor may be influencing the relationship between the variables.
EDA analysis cannot definitively determine causality between the variables, only show that there is a relationship but not whether it is causal or not, they may both increase at the same time due to the influence of a third variable that is not being measured.
